Gaussian variance estimation in simulated data sets

### 2018-11-09 analysis/gaussvarest.Rmd
wflow_publish("gaussvarest.Rmd")

# ---
# title: "Gaussian variance estimation in simulated data sets"
# ---
#
# This script implements the "Gaussian variance estimation" simulation
# experiments in the paper. In particular, we compare the Mean Field
# Variational Bayes method (MFVB) against SMASH in two scenarios. See
# the paper for the full description of these scenarios.
#
# The figure and table generated at the end of this script should
# match up with the figure and table shown in the paper.
#
# Note that this script could take an hour or several hours to
# complete as it runs the two methods on 100 simulated data sets for
# each of the two scenarios.
#
# We thank Menictas & Wand for generously sharing code that was used
# to implement these experiments.
#

# -----
# TO DO: Remove steps that are not included in the paper.
# -----

# SET UP ENVIRONMENT
# ------------------
# Load the smashr package, and some functions used in the analysis.
library(smashr)
source("../code/mfvb.functions.R")

# SCRIPT PARAMETERS
# -----------------
# Number of data sets simulated in the first (nsim1) and second
# (nsim2) simulation scenarios.
nsim1 <- 100
nsim2 <- 100

# Set hyperparameters.
Au.hyp      <- 1e5
Av.hyp      <- 1e5
sigsq.gamma <- 1e10
sigsq.beta  <- 1e10

# Set colours.
mainCol <- "darkslateblue"
ptCol   <- "paleturquoise3"
lineCol <- "skyblue"
axisCol <- "black"

# Set plotting size parameters.
cex.pt      <- 0.75
cex.mainVal <- 1.7
cex.labVal  <- 1.3
xlabVal     <- "x"

# FIRST SIMULATION SCENARIO
# -------------------------
mse.mu.uneven.mfvb <- 0
mse.mu.uneven.haar <- 0
mse.mu.uneven.s8   <- 0

@@

  mse.mu.uneven.s8[j]=mean((mu.est.inter-fTrue(xgOrig))^2)
  mse.sd.uneven.s8[j]=mean((sqrt(var.est.inter)-exp((loggTrue(xgOrig))/2))^2)
  mse.sd.uneven.j[j] =mean((sqrt(var.est.inter.j)-exp((loggTrue(xgOrig))/2))^2)
  mse.sd.uneven.s8.j[j] =
    mean((sqrt(var.est.inter.s8.j) - exp((loggTrue(xgOrig))/2))^2)
  lmse.sd.uneven.s8[j]=mean((log(sqrt(var.est.inter))-(loggTrue(xgOrig))/2)^2)
  lmse.sd.uneven.j[j]=mean((log(sqrt(var.est.inter.j))-(loggTrue(xgOrig))/2)^2)
  lmse.sd.uneven.s8.j[j] =
    mean((log(sqrt(var.est.inter.s8.j))-(loggTrue(xgOrig))/2)^2)
}
cat("\n")

# SECOND SIMULATION SCENARIO
# --------------------------
mse.mu.even.mfvb <- 0
mse.mu.even.haar <- 0
mse.mu.even.s8   <- 0

@@

# SUMMARIZE RESULTS
# -----------------
# Compare this table against Table 1 in the paper.
cat("Summarizing results of simulations.\n")
cat(sprintf("MSE averaged across %d simulations in Scenario 1,\n",nsim1))
cat(sprintf("and averaged across %d simulations in Scenario 2:\n",nsim2))
mse.table <- rbind(c(mean(mse.mu.uneven.mfvb),mean(mse.sd.uneven.mfvb),
                     mean(mse.mu.even.mfvb),mean(mse.sd.even.mfvb)),
                   c(mean(mse.mu.uneven.s8),mean(mse.sd.uneven.s8),
                     mean(mse.mu.even.s8),mean(mse.sd.even.s8)))
rownames(mse.table) <- c("MFVB","SMASH")
colnames(mse.table) <- c("mean","sd","mean","sd")
cat("         Scenario 1      Scenario 2   \n")
cat("      --------------- ----------------\n")
print(mse.table)

>>>>

---
title: "Gaussian variance estimation in simulated data sets"
author: "Zhengrong Xing, Peter Carbonetto and Matthew Stephens"
output: workflowr::wflow_html
---

This script implements the "Gaussian variance estimation" simulation
experiments in the paper. In particular, we compare the Mean Field
Variational Bayes method (MFVB) against SMASH in two scenarios.

The figure and table generated at the end of this script should
match up with the figure and table shown in the paper.

Running the code could take several hours to complete as it runs the
two methods on 100 simulated data sets for each of the two scenarios.

We thank M. Menictas & M. Wand for generously sharing code that was
used to implement these experiments.

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,fig.align = "center")
```

Initial setup instructions
--------------------------

To run this example on your own computer, please follow these setup
instructions. These instructions assume you already have R and/or
RStudio installed on your computer.

First, download or clone the [git repository][smash-github] on your
computer.

Launch R, and change the working directory to be the "analysis" folder
inside your local copy of the git repository.

Finally, install the smashr package from GitHub:

```{r install-pkgs-2, eval=FALSE}
devtools::install_github("stephenslab/smashr",upgrade_dependencies = FALSE)
```

See the "Session Info" at the bottom for the versions of the software
and R packages that were used to generate the results shown below.

Set up R environment
--------------------

Load the smashr package, as well as some functions used in the
analysis below.

```{r load-pkgs}
library(smashr)
source("../code/mfvb.R")
```

Analysis settings
-----------------

Specify the number of data sets simulated in the first and second
simulation scenarios.

```{r analysis-settings-1}
nsim1 <- 10
nsim2 <- 10
```

Next, specify the hyperparameters used in running the MFVB method.

```{r analysis-settings-2}
Au.hyp      <- 1e5
Av.hyp      <- 1e5
sigsq.gamma <- 1e10
sigsq.beta  <- 1e10
```

These variables specify some colours used in the plots.

```{r analysis-settings-3}
mainCol <- "darkslateblue"
ptCol   <- "paleturquoise3"
lineCol <- "skyblue"
axisCol <- "black"
```

These are additional plotting parameters.

```{r analysis-settings-4}
cex.pt      <- 0.75
cex.mainVal <- 1.7
cex.labVal  <- 1.3
xlabVal     <- "x"
```

First simulation scenario
-------------------------

```{r simulation-1}
mse.mu.uneven.mfvb <- 0
mse.mu.uneven.haar <- 0
mse.mu.uneven.s8   <- 0

@@

  mse.mu.uneven.s8[j]=mean((mu.est.inter-fTrue(xgOrig))^2)
  mse.sd.uneven.s8[j]=mean((sqrt(var.est.inter)-exp((loggTrue(xgOrig))/2))^2)
  mse.sd.uneven.j[j] =mean((sqrt(var.est.inter.j)-exp((loggTrue(xgOrig))/2))^2)
  mse.sd.uneven.s8.j[j] =
    mean((sqrt(var.est.inter.s8.j) - exp((loggTrue(xgOrig))/2))^2)
  lmse.sd.uneven.s8[j]=mean((log(sqrt(var.est.inter))-(loggTrue(xgOrig))/2)^2)
  lmse.sd.uneven.j[j]=mean((log(sqrt(var.est.inter.j))-(loggTrue(xgOrig))/2)^2)
  lmse.sd.uneven.s8.j[j] =
    mean((log(sqrt(var.est.inter.s8.j))-(loggTrue(xgOrig))/2)^2)
}
cat("\n")
```

```
# SECOND SIMULATION SCENARIO
# --------------------------
mse.mu.even.mfvb <- 0
mse.mu.even.haar <- 0
mse.mu.even.s8   <- 0

@@

# SUMMARIZE RESULTS
# -----------------
# Compare this table against Table 1 in the paper.
cat("Summarizing results of simulations.\n")
cat(sprintf("MSE averaged across %d simulations in Scenario 1,\n",nsim1))
cat(sprintf("and averaged across %d simulations in Scenario 2:\n",nsim2))
mse.table <- rbind(c(mean(mse.mu.uneven.mfvb),mean(mse.sd.uneven.mfvb),
                     mean(mse.mu.even.mfvb),mean(mse.sd.even.mfvb)),
                   c(mean(mse.mu.uneven.s8),mean(mse.sd.uneven.s8),
                     mean(mse.mu.even.s8),mean(mse.sd.even.s8)))
rownames(mse.table) <- c("MFVB","SMASH")
colnames(mse.table) <- c("mean","sd","mean","sd")
cat("         Scenario 1      Scenario 2   \n")
cat("      --------------- ----------------\n")
print(mse.table)
```
